### 2025-09-18 paper/sec/introduction.tex
Merge 1.1 into 1.

\section{Introduction}

\begin{itemize}
    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
    \item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
    \item We see this in peer review, misinformation, retracted papers, …
\end{itemize}

\rpnote{Small example here}

@@

Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed
integrating interactive ``provenance queries'' directly into charts and figures. These interactive queries
allow readers to interact with visual elements to see the data relevant to those elements. The key feature of
these approaches is that the interactions are provided automatically, via \emph{transparent programming
languages} (either query languages or general-purpose languages) which track dependencies as data flows
through a computation.

>>>>

\section{Introduction: Transparent Documents}

Making sense of data-driven claims is hard, even if the evidence base (data sets and related analyses) is
open. We see this in peer review, misinformation, retracted papers, \rpnote{supporting citations}.
\rpnote{Small example here?}

Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed more
transparent charts and figures by integrating interactive ``provenance queries'' directly into visual outputs.
These interactive queries allow readers to interact with visual elements to see the data relevant to those
elements. The key feature of these approaches is that the interactions are provided automatically, via
\emph{transparent programming languages} (either query languages or general-purpose languages) which track
dependencies as data flows through a computation.

\begin{itemize}
    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
\end{itemize}

@@
